Uncertainty propagation in deep neural networks for sparse coding

### 2018-03-10 paper/BayesianLISTA.tex
minor fix
word check

\begin{abstract}
In the Bayesian neural network framework prior distributions are imposed on weights of a network and a posterior distribution of an output is inferred, rather than a point prediction provided by conventional neural networks. Therefore, Bayesian neural networks estimate both an output for a test data and uncertainty of this prediction. Uncertainty of the prediction is a result of propagating weight uncertainty through layers of a network. We propose a novel method of propagating uncertainty through the soft thresholding nonlinearity. The method expands the area of Bayesian neural networks and enables the soft thresholding nonlinearity to be included in Bayesian deep neural networks for the first time. As an example we design a Bayesian Learned Iterative Shrinkage-Thresdholding Algorithm (Bayesian \textsc{lista}) network. We develop a highly efficient posterior inference algorithm based on probabilistic backpropagation. A series of experiments on sparse coding shows that the proposed framework provide both accurate predictions and quantified estimates of uncertainty in these predictions.
\end{abstract}

@@

\section{\uppercase{Uncertainty propagation through soft thresholding}}
\label{sec:fprop}
For every layer of the Baysian \textsc{lista} we assume that $\widehat{\boldsymbol\beta}_{l-1}$ has the spike and slab distribution with parameters: $\boldsymbol\omega$ is a probability of a spike, $\mathbf{m}$ is a mean of a slab Gaussian distribution, and $\mathbf{v}$ is a variance of the slab distribution:
\begin{equation}
[\widehat{\boldsymbol\beta}_{l-1}]_d \sim \omega_d \delta_0 + (1 - \omega_d)\mathcal{N}(m_d, v_d),
\end{equation}
where $\omega_d$, $m_d$, and $v_d$ are components of vectors $\boldsymbol\omega$, $\mathbf{m}$, and $\mathbf{v}$, respectively, $\delta_0$ is the delta-function that represents a spike. 
We initialise $\widehat{\boldsymbol\beta}_{0}$ with the spike and slab distribution with parameters $\boldsymbol\omega = \mathbf{0}$, $\mathbf{m} = \mathbf{0}$, and $\mathbf{v} = \mathbf{1}$.

@@

\subsection{\uppercase{Main layers}}
The distributions of the inputs at the step (\ref{eq:l_dense_layer}) of the Baysian \textsc{lista} for each $l$ are: the vector $\mathbf{b}$ and matrix $\mathbf{S}$ consist of Gaussian-distributed components and $\widehat{\boldsymbol\beta}_{l-1}$ is a vector of the spike and slab random variables.

@@

Let $\mathbf{e}_l = \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$ at the step (\ref{eq:l_dense_layer}) of the Baysian \textsc{lista}. We assume that $\mathbf{S}$ and $\widehat{\boldsymbol\beta}_{l-1}$ are mutually independent. Then according to the Central Limit Theorem $[ \mathbf{e}_l ]_d$ can be approximated as a Gaussian-distributed variable when $D$ is sufficiently large. The parameters of the of the approximating Gaussian distribution are set to the marginal mean and variance given in Lemma~\ref{thm:matrix_vector}. The quality of this approximation is discussed in section \ref{sec:approx_quality}.

@@

Then $\widehat{\boldsymbol\beta}_{l}$ at the step (\ref{eq:l_thr}) of the Baysian \textsc{lista} is the result of soft thresholding of a Gaussian distribution, which is approximated with the spike and slab distribution,  similar to the step (\ref{eq:thr_first}).

@@

Parameters of the approximating posterior distribution are then updated with the derivates of this normalisation constant. [ADD ABOUT RECURRENT ESTIMATION]

@@

Experiments on the synthetic and \textsc{mnist} datasets demonstrate that the proposed algorithm preserves the predictive accuracy of non-Bayesian methods while also providing posterior estimates. We also show that when the training data is very small the proposed algorithm significantly outperforms the classical \textsc{lista} in terms of predictive accuracy. Experiments on active learning demonstrate that the proposed Baysian \textsc{lista} gives accurate posterior estimates that can be used for selection of a next data point where a label should be obtain for.

The shrinkage parameter $\lambda$ is currently treated as a deterministic hyperparameter for the Baysian \textsc{lista}. In future we plan to incorporate it into the model treating it as a random variable. For this we need to extend both the uncertainty propagation method to include uncertainty of $\lambda$ and the probabilistic backpropagation algorithm to estimate its posterior. We will study stochastic expectation propagation applicability in order to replace ADF that may improve further the quality of posterior estimates~\citep{li2015stochastic}.    
\bibliography{bibliography}
\bibliographystyle{icml2018}

>>>>

\begin{abstract}
In the Bayesian neural network framework prior distributions are imposed on weights of a network and a posterior distribution of an output is inferred, rather than a point prediction provided by conventional neural networks. Therefore, Bayesian neural networks estimate both an output for a test data and uncertainty of this prediction. Uncertainty of the prediction is a result of weight uncertainty propagation through the layers of a network. We propose a novel method of propagating uncertainty through the soft thresholding nonlinearity. The method expands the area of Bayesian neural networks and enables the soft thresholding nonlinearity to be included in Bayesian deep neural networks for the first time. As an example we design a Bayesian Learned Iterative Shrinkage-Thresdholding Algorithm (Bayesian \textsc{lista}) network. We develop a highly efficient posterior inference algorithm based on probabilistic backpropagation. A series of experiments on sparse coding shows that the proposed framework provides both accurate predictions and quantified estimates of uncertainty in these predictions.
\end{abstract}

@@

\section{\uppercase{Uncertainty propagation through soft thresholding}}
\label{sec:fprop}
For every layer of the Bayesian \textsc{lista} we assume that $\widehat{\boldsymbol\beta}_{l-1}$ has the spike and slab distribution with parameters: $\boldsymbol\omega$ is a probability of a spike, $\mathbf{m}$ is a mean of a slab Gaussian distribution, and $\mathbf{v}$ is a variance of the slab distribution:
\begin{equation}
[\widehat{\boldsymbol\beta}_{l-1}]_d \sim \omega_d \delta_0 + (1 - \omega_d)\mathcal{N}(m_d, v_d),
\end{equation}
where $\omega_d$, $m_d$, and $v_d$ are the components of vectors $\boldsymbol\omega$, $\mathbf{m}$, and $\mathbf{v}$, respectively, $\delta_0$ is the delta-function that represents a spike. 
We initialise $\widehat{\boldsymbol\beta}_{0}$ with the spike and slab distribution with parameters $\boldsymbol\omega = \mathbf{0}$, $\mathbf{m} = \mathbf{0}$, and $\mathbf{v} = \mathbf{1}$.

@@

\subsection{\uppercase{Main layers}}
The distributions of the inputs at the step (\ref{eq:l_dense_layer}) of the Bayesian \textsc{lista} for each $l$ are: the vector $\mathbf{b}$ and matrix $\mathbf{S}$ consist of Gaussian-distributed components and $\widehat{\boldsymbol\beta}_{l-1}$ is a vector of the spike and slab random variables.

@@

Let $\mathbf{e}_l = \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$ at the step (\ref{eq:l_dense_layer}) of the Bayesian \textsc{lista}. We assume that $\mathbf{S}$ and $\widehat{\boldsymbol\beta}_{l-1}$ are mutually independent. Then according to the Central Limit Theorem $[ \mathbf{e}_l ]_d$ can be approximated as a Gaussian-distributed variable when $D$ is sufficiently large. The parameters of the of the approximating Gaussian distribution are set to the marginal mean and variance given in Lemma~\ref{thm:matrix_vector}. The quality of this approximation is discussed in section \ref{sec:approx_quality}.

@@

Then $\widehat{\boldsymbol\beta}_{l}$ at the step (\ref{eq:l_thr}) of the Bayesian \textsc{lista} is the result of soft thresholding of a Gaussian distribution, which is approximated with the spike and slab distribution,  similar to the step (\ref{eq:thr_first}).

@@

Parameters of the approximating posterior distribution are then updated with the derivatives of this normalisation constant.

@@

Experiments on the synthetic and \textsc{mnist} datasets demonstrate that the proposed algorithm preserves the predictive accuracy of non-Bayesian methods while also providing posterior estimates. We also show that when the training data is very small the proposed algorithm significantly outperforms the classical \textsc{lista} in terms of predictive accuracy. Experiments on active learning demonstrate that the proposed Bayesian \textsc{lista} gives accurate posterior estimates that can be used for selection of a next data point where a label should be obtain for.

The shrinkage parameter $\lambda$ is currently treated as a deterministic hyperparameter for the Bayesian \textsc{lista}. In future we plan to incorporate it into the model treating it as a random variable. For this we need to extend both the uncertainty propagation method to include uncertainty of $\lambda$ and the probabilistic backpropagation algorithm to estimate its posterior. We will study stochastic expectation propagation applicability in order to replace ADF that may improve further the quality of posterior estimates~\citep{li2015stochastic}.    
\bibliography{bibliography}
\bibliographystyle{icml2018}
